---
title: "Computational Modeling - Assignment 1"
author: "Riccardo Fusaroli"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. this markdown has 2 parts as it spans 2 weeks of teaching

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

# 1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
## - First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r setup, include = TRUE}
pacman::p_load(tidyverse, rethinking)
```

### Manually first for overview - making a function after
```{r}
# Defining things
dens <- 101
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

# Test prior - does it look crazy?
dens(rbinom(1e4, 9, runif(1e4, 0, 1))) 

# Computing the likelihood at each value in grid
likelihood <- dbinom(3 , size = 6 , prob = p_grid) #creates a binomial distribution using that he answered 6 out of 9. 

# Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

# Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

# Making a dataframe and making the plot
d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Knowledge of CogSci")+ ylab("posterior probability")

# Calculating the probability of him answering above 50% 
#sum(d$posterior[which(d$grid == 0.5):101])/sum(d$posterior)
p_over_0.5 <- sum(d$posterior[p_grid > 0.5]) #NB I do not divide by the sum here. Reason? Hasnt it already been done? Implement if needed
p_over_0.5 #0.48 - før fik vi sådan noget 51, så det virker som om der er noget der bliver udeladt her. Betydning? Evt sæt den til => i stedet for >?
```
## - Then implement a quadratic approximation (hint check paragraph 2.4.2!). N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
# Using quap first to do quadratic approximation
globe.qa <- quap(
  alist(
    W~dbinom(C+W,p),#binomial likelihood
    p~dunif(0,1) #uniform prior
  ), data = list(C=3, W=3))

# Display summary of quadratic approximation
precis(globe.qa)

# Analytical calculation and plot of quadratic approximation
W <- 3
C <- 3
curve(dbeta(x,W+1,C+1), from = 0, to = 1)

# Quadratic approximation with posterior mean value of 0.5 and curvature of 0.2.
curve(dnorm(x,0.5,0.2), lty = 2, add = TRUE) 

# Getting values
p <- curve(dnorm(x,0.5,0.2), lty = 2, add = TRUE)

# BEFORE IT SAID: area_above_50 <- sum(p$y[which(p$x == 0.5)]:length(p$y))/sum(p$y)
# Calculating the probability of answering above chance
p_over_0.5 <- sum(p$y[p$x > 0.5]) #NB I do not divide by the sum here. Reason? Hasnt it already been done? Implement if needed
p_over_0.5 #48, ikke 0.489 NB som ovenover
```
We have a posterior mean value of p = 0.5, it calls it the Mean. Curvature is the sd, of 0.2. Assuming the posterior is Gaussian, it is maximized at 0.50, and its standard deviation is 0.20. 

# 2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.

### Making a function for finding the posterior probability + area above 50% - but it can be sooo heavily optimized
NB: *pre-specify part can be implemented in function but since it will prob be equal in the next points in the assignments I thought meh for now*
NB: *doesnt have the check of priors, the rbinom one - prob important, say something about it / use it before the function*
```{r}
# Pre-specify what you want these two values to be before using the function on something
dens = 202
p_grid <- seq(from = 0 ,
                to = 1 ,
                length.out = dens)

# Function returning DF and PLOT
FindMyPosteriors <- function(correct, total, prior, p_grid) {
  likelihood <- dbinom(correct , size = total , prob = p_grid)
  unstd.posterior <- likelihood*prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  d <- data.frame(
    grid = p_grid,
    posterior = posterior,
    prior = prior,
    likelihood = likelihood)
  
  p_over_0.5 <- sum(d$posterior[p_grid > 0.5]) #NB I do not divide by the sum here. Reason? Hasnt it already been done? Implement if needed

  df = data.frame(p_grid,unstd.posterior,likelihood,posterior,prior,p_over_0.5)
   plot <-
    ggplot(d, aes(grid, posterior)) + geom_point() + geom_line() + theme_classic() + geom_line(aes(grid, prior /dens), color = 'red') + xlab("Knowledge of CogSci") + ylab("posterior probability") + ggtitle("Plot from function, evt gør noget med paste her med navnene og et loop der går igennem lærerne?")
   
  return(list(df,plot))
}

# Two tests to see what you will get: access df in object[[1]] and access plot in object[[2]]
t_norm <- FindMyPosteriors(correct = 6,total = 12,prior = dnorm(p_grid, 0.8, 0.2), p_grid)
t_uniform <- FindMyPosteriors(correct = 6,total = 12,prior <- rep(1 , dens), p_grid)

t_uniform[[1]] #df
t_uniform[[2]] #plot

t_norm[[1]] #df
t_norm[[2]] #plot

```

# 3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?

###  Using function to find teachers smarter knowledge
```{r}
# Information equal for all teachers
dens = 101
p_grid = seq(from = 0 , to = 1 , length.out = dens)
prior = dnorm(p_grid, 0.8,0.2)
dens(rbinom(1e4, 9, runif(1e4,0,1))) # Checking prior, not sure I understand the numbers here, the 9 fx? wat are you
```


```{r}
# Riccardo: 3 correct out of 6 total 
SmarterRiccardo <- FindMyPosteriors(correct = 3,total = 6, prior = prior, p_grid = p_grid)
SmarterRiccardo[[1]] # 0.83 - Is it right? So that's 83% probability summed up that Riccardos knowledge of CogSci is over 50%? Formulation of this answer instead: is it the same as saying there is 83% probability taking both prior and data into account that Riccardo answers more right than chance level? 
SmarterRiccardo[[2]]+ggtitle("Smarter Riccardo's knowledge of CogSci") # I want a line put at the above 50 level, preferably that it fills out the area

# Kristian: 2 correct out of 2 total (gets bored)
SmarterTylén <- FindMyPosteriors(correct = 2, total = 2, prior = prior, p_grid = p_grid)
SmarterTylén[[1]] # p_over_5: 0.97. Is it right? Makes sense? 
SmarterTylén[[2]]+ggtitle("Smarter Tylén's knowledge of CogSci")

# Daina: 160 correct out of 198 questions (never gets bored)
SmarterDaina <- FindMyPosteriors(correct = 160, total = 198, prior = prior, p_grid = p_grid)
SmarterDaina[[1]] # p_over 50 = 1
SmarterDaina[[2]]+ggtitle("Smarter Daina's knowledge of CogSci") #wow

# Mikkel: 66 correct answers out of 132 questions (Mikkel: He's just like us)
SmarterMikkel <- FindMyPosteriors(correct = 66, total = 132, prior = prior, p_grid = p_grid)
SmarterMikkel[[1]] #p_over_50 = 0.57
SmarterMikkel[[2]]+ggtitle("Smarter Mikkel's knowledge of CogSci")
```

## 3a. Produce plots of the prior and posterior for each teacher.
Already been done a bit above, but here is a grid: 
```{r}
# Collecting the plots in a PlotGrid
Smarter_teachers <-
  cowplot::plot_grid(
    SmarterRiccardo[[2]] + ggtitle("Smarter Riccardo"),
    SmarterTylén[[2]] + ggtitle("Smarter Tylén"),
    SmarterDaina[[2]] + ggtitle("Smarter Daina"),
    SmarterMikkel[[2]] + ggtitle("Smarter Mikkel")
  )
Smarter_teachers

```

# 4 . You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}
# Uniform prior
dens = 101 # arbitrary number - what do we wanna put? 
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

Riccardo_uf <- FindMyPosteriors(correct = 300,total = 600, prior = prior, p_grid = p_grid)
Riccardo_uf[[1]]
Riccardo_uf[[2]]+ggtitle("Uniform prior - Riccardo's knowledge of CogSci") 

Tylén_uf <- FindMyPosteriors(correct = 200, total = 200, prior = prior, p_grid = p_grid)
Tylén_uf[[1]] 
Tylén_uf[[2]]+ggtitle("Uniform prior - Tylén's knowledge of CogSci")

Daina_uf <- FindMyPosteriors(correct = 16000, total = 19800, prior = prior, p_grid = p_grid)
Daina_uf[[1]] 
Daina_uf[[2]]+ggtitle("Uniform prior - Daina's knowledge of CogSci") 

Mikkel_uf <- FindMyPosteriors(correct = 6600, total = 13200, prior = prior, p_grid = p_grid)
Mikkel_uf[[1]] 
Mikkel_uf[[2]]+ggtitle("Uniform prior - Mikkel's knowledge of CogSci")

# Collecting the plots in a PlotGrid
uniform_teachers <-
  cowplot::plot_grid(
    Riccardo_uf[[2]] + ggtitle("Uni. prior, big data - Riccardo"),
    Tylén_uf[[2]] + ggtitle("Uni. prior, big data - Tylén"),
    Daina_uf[[2]] + ggtitle("Uni. prior, big data - Daina"),
    Mikkel_uf[[2]] + ggtitle("Uni. prior, big data - Mikkel")
  )
uniform_teachers

# Making them with regular data and uniform priors to compare
dens = 101 
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

Riccardo_uf_reg <- FindMyPosteriors(correct = 3,total = 6, prior = prior, p_grid = p_grid)
Riccardo_uf_reg[[1]]
Riccardo_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Riccardo") 

Tylén_uf_reg <- FindMyPosteriors(correct = 2, total = 2, prior = prior, p_grid = p_grid)
Tylén_uf_reg[[1]] 
Tylén_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Tylén")

Daina_uf_reg <- FindMyPosteriors(correct = 160, total = 198, prior = prior, p_grid = p_grid)
Daina_uf_reg[[1]] 
Daina_uf_reg[[2]]+ggtitle("Uniform prior,  reg. data - Daina") 

Mikkel_uf_reg <- FindMyPosteriors(correct = 66, total = 132, prior = prior, p_grid = p_grid)
Mikkel_uf_reg[[1]] 
Mikkel_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Mikkel")

uniform_reg_teachers <-
  cowplot::plot_grid(
    Riccardo_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Riccardo"),
    Tylén_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Tylén"),
    Daina_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Daina"),
    Mikkel_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Mikkel")
  )
uniform_reg_teachers

# Compare
cowplot::plot_grid(uniform_teachers,uniform_reg_teachers)
```
5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r}

```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)? 
```{r}

```

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Daina: 148 correct answers out of 172 questions (again, Daina never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)








##### 
```{r}
# # OLD OLD OLD OLD OLD BUT DONT DELETE JUST YET
# # Equal for all: dens <- 101
# p_grid <- seq(from = 0 , to = 1 , length.out = dens)
# prior <- rep(1 , dens)
# dens(rbinom(1e4, 9, runif(1e4, 0, 1))) 

# # Riccardo: 3 correct answers out of 6 questions
# likelihood <- dbinom(3 , size = 6 , prob = p_grid) #creates a binomial distribution using that he answered 6 out of 9. 
# unstd.posterior <- likelihood * prior
# posterior <- unstd.posterior / sum(unstd.posterior)
# d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
# smarter_riccardo <- ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Smarter Riccardo's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("AMAZING 4-PLOTS - BUT NOW")
# 
# # Calculating the probability of him answering above 50%. First looking for 50% mark, then summing up from there to find area under the curve: 
# sum(d$posterior[which(d$grid == 0.5):101])/sum(d$posterior)
# 
# # Kristian: 2 correct answers out of 2 questions (then he gets bored)
# likelihood <- dbinom(2 , size = 2 , prob = p_grid) 
# unstd.posterior <- likelihood * prior
# posterior <- unstd.posterior / sum(unstd.posterior)
# d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
# smarter_tylen <- ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Smarter Tylén's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("THEY'RE SMARTER (WE ASSUME...")
# 
# # Calculating the probability of him answering above 50%. First looking for 50% mark, then summing up from there to find area under the curve: 
# sum(d$posterior[which(d$grid == 0.5):101])/sum(d$posterior)
# 
# # Daina: 160 correct answers out of 198 questions (Daina never gets bored)
# likelihood <- dbinom(160 , size = 198 , prob = p_grid)
# unstd.posterior <- likelihood * prior
# posterior <- unstd.posterior / sum(unstd.posterior)
# d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
# smarter_daina <- ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Smarter Daina's knowledge of CogSci")+ ylab("posterior probability")
# 
# # Calculating the probability of him answering above 50%. First looking for 50% mark, then summing up from there to find area under the curve: 
# sum(d$posterior[which(d$grid == 0.5):101])/sum(d$posterior)
# 
# # Mikkel: 66 correct answers out of 132 questions
# likelihood <- dbinom(66 , size = 132 , prob = p_grid)
# unstd.posterior <- likelihood * prior
# posterior <- unstd.posterior / sum(unstd.posterior)
# d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)
# smarter_mikkel <- ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Smarter Mikkel's knowledge of CogSci")+ ylab("posterior probability")
# 
# # Calculating the probability of him answering above 50%. First looking for 50% mark, then summing up from there to find area under the curve: 
# sum(d$posterior[which(d$grid == 0.5):101])/sum(d$posterior)
# 
# # PlotGrid
# smarter_teachers <- cowplot::plot_grid(smarter_riccardo,smarter_tylen,smarter_daina,smarter_mikkel)
# smarter_teachers
# ```
# ```{r}
# # Comparing prior_dumb and prior_smart
# teachers
# smarter_teachers
```
