---
title: "Computational Modeling - Assignment 1"
author: "Riccardo Fusaroli"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. this markdown has 2 parts as it spans 2 weeks of teaching

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

# 1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
## - First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r setup, include = TRUE}
pacman::p_load(tidyverse, rethinking)
```

### Doing it manually first for overview - making a function after
```{r}
# Defining things
dens <- 101
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

# Test prior - does it look crazy?
dens(rbinom(1e4, 9, runif(1e4, 0, 1))) 

# Computing the likelihood at each value in grid
likelihood <- dbinom(3 , size = 6 , prob = p_grid) #creates a binomial distribution using that he answered 6 out of 9. 

# Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

# Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

# Making a data frame and making the plot
d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

ggplot(d, aes(grid,posterior)) + geom_point() +geom_line()+theme_classic()+ geom_line(aes(grid, prior/dens),color= 'red')+ xlab("Knowledge of CogSci")+ ylab("posterior probability") + ggtitle("Posterior distribution - CogSci smartness")

# Calculating the probability of him answering above 50%. This is an interval of defined boundary. To calculate that we need to simply count up the samples that satisfy the criterion, and then divide with the number of samples.
p_over_0.5 <- sum(d$posterior[p_grid > 0.5])/sum(d$posterior)

p_over_0.5
```
## - Then implement a quadratic approximation (hint check paragraph 2.4.2!). N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
# Using quap first to do quadratic approximation
globe.qa <- quap(
  alist(
    W~dbinom(C+W,p),#binomial likelihood
    p~dunif(0,1) #uniform prior
  ), data = list(C=3, W=3))

# Display summary of quadratic approximation
precis(globe.qa)

# Analytical calculation and plot of quadratic approximation
W <- 3
C <- 3
curve(dbeta(x,W+1,C+1), from = 0, to = 1)

# Quadratic approximation with posterior mean value of 0.5 and curvature of 0.2.
curve(dnorm(x,0.5,0.2), lty = 2, add = TRUE) 

# Getting values
p <- curve(dnorm(x,0.5,0.2), lty = 2, add = TRUE)

# Calculating the probability of answering above chance
p_over_0.5 <- sum(p$y[p$x > 0.5])/sum(p$y)

p_over_0.5 
```
We have a posterior mean value of p = 0.5, it calls it the mean. Curvature is the sd, of 0.2. Assuming the posterior is Gaussian, it is maximized at 0.50, and its standard deviation is 0.20. 

# 2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
### Making a function for finding the posterior probability + area above 50%
NB: *pre-specify part can be implemented in function but since it will prob be equal in the next points in the assignments I thought meh for now*
NB: *doesnt have the check of priors, the rbinom one - prob important, say something about it / use it before the function*
```{r}
# Pre-specify what you want these two values to be before using the function on something
dens = 202
p_grid <- seq(from = 0 ,
                to = 1 ,
                length.out = dens)

# Function returning DF and PLOT
FindMyPosteriors <- function(correct, total, prior, p_grid) {
  likelihood <- dbinom(correct , size = total , prob = p_grid)
  unstd.posterior <- likelihood*prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  d <- data.frame(
    grid = p_grid,
    posterior = posterior,
    prior = prior,
    likelihood = likelihood)
  
  p_over_0.5 <- sum(d$posterior[p_grid > 0.5])/sum(d$posterior)

  df = data.frame(p_grid,unstd.posterior,likelihood,posterior,prior,p_over_0.5)
   plot <-
    ggplot(d, aes(grid, posterior)) + geom_point() + geom_line() + theme_classic() + geom_line(aes(grid, prior /dens), color = 'red') + xlab("Knowledge of CogSci") + ylab("posterior probability") + ggtitle("Plot from function") # IDE:  evt gør noget med paste her med navnene og et loop der går igennem lærerne?
   
  return(list(df,plot))
}

# Two tests to see what the function returns: access df in object[[1]] and access plot in object[[2]]
t_norm <- FindMyPosteriors(correct = 6,total = 12,prior = dnorm(p_grid, 0.8, 0.2), p_grid)
t_uniform <- FindMyPosteriors(correct = 6,total = 12,prior <- rep(1 , dens), p_grid)

t_uniform[[1]] #df
t_uniform[[2]] #plot

t_norm[[1]] #df
t_norm[[2]] #plot

```

# 3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?

###  Using function to find teachers smarter knowledge
```{r}
# Information equal for all teachers
dens = 101
p_grid = seq(from = 0 , to = 1 , length.out = dens)
prior = dnorm(p_grid, 0.8,0.2)
dens(rbinom(1e4, 9, runif(1e4,0,1))) # Checking prior, not sure I understand the numbers here, the 9 fx? 9 tosses in the book where we use rbinom to do a posterior predictive check. #simplehist(rbinom(1e4, size = 9, prob = ???)
```

*NB: HOW TO INTERPRET THE THINGS BELOW?*
E.g. For Riccardo: 
It returns a posterior probability of 0.83. That means that what? That there is 84% probability that Riccardos knowledge is over 50%. That is not what is asked though is it? Dont they just ask who is best? 
```{r}
# Riccardo: 3 correct out of 6 total 
SmarterRiccardo <- FindMyPosteriors(correct = 3,total = 6, prior = prior, p_grid = p_grid)
SmarterRiccardo[[1]] 
SmarterRiccardo[[2]]+ggtitle("Smarter Riccardo's knowledge of CogSci") 

# Kristian: 2 correct out of 2 total (gets bored)
SmarterTylén <- FindMyPosteriors(correct = 2, total = 2, prior = prior, p_grid = p_grid)
SmarterTylén[[1]] 
SmarterTylén[[2]]+ggtitle("Smarter Tylén's knowledge of CogSci")

# Daina: 160 correct out of 198 questions (never gets bored)
SmarterDaina <- FindMyPosteriors(correct = 160, total = 198, prior = prior, p_grid = p_grid)
SmarterDaina[[1]] 
SmarterDaina[[2]]+ggtitle("Smarter Daina's knowledge of CogSci") #wow

# Mikkel: 66 correct answers out of 132 questions (Mikkel: He's just like us)
SmarterMikkel <- FindMyPosteriors(correct = 66, total = 132, prior = prior, p_grid = p_grid)
SmarterMikkel[[1]] 
SmarterMikkel[[2]]+ggtitle("Smarter Mikkel's knowledge of CogSci")
```

## 3a. Produce plots of the prior and posterior for each teacher.
Already been done a bit above, but here is a grid: 
```{r}
# Collecting the plots in a grid
Smarter_teachers <-
  cowplot::plot_grid(
    SmarterRiccardo[[2]] + ggtitle("Smarter Riccardo"),
    SmarterTylén[[2]] + ggtitle("Smarter Tylén"),
    SmarterDaina[[2]] + ggtitle("Smarter Daina"),
    SmarterMikkel[[2]] + ggtitle("Smarter Mikkel")
  )
Smarter_teachers

```

# 4 . You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}
# Uniform prior and a lot of data
dens = 101 
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

Riccardo_uf <- FindMyPosteriors(correct = 300,total = 600, prior = prior, p_grid = p_grid)
Riccardo_uf[[1]]
Riccardo_uf[[2]]+ggtitle("Uniform prior - Riccardo's knowledge of CogSci") 

Tylén_uf <- FindMyPosteriors(correct = 200, total = 200, prior = prior, p_grid = p_grid)
Tylén_uf[[1]] 
Tylén_uf[[2]]+ggtitle("Uniform prior - Tylén's knowledge of CogSci")

Daina_uf <- FindMyPosteriors(correct = 16000, total = 19800, prior = prior, p_grid = p_grid)
Daina_uf[[1]] 
Daina_uf[[2]]+ggtitle("Uniform prior - Daina's knowledge of CogSci") 

Mikkel_uf <- FindMyPosteriors(correct = 6600, total = 13200, prior = prior, p_grid = p_grid)
Mikkel_uf[[1]] 
Mikkel_uf[[2]]+ggtitle("Uniform prior - Mikkel's knowledge of CogSci")

# Collecting the plots in a PlotGrid
uniform_teachers <-
  cowplot::plot_grid(
    Riccardo_uf[[2]] + ggtitle("Uni. prior, big data - Riccardo"),
    Tylén_uf[[2]] + ggtitle("Uni. prior, big data - Tylén"),
    Daina_uf[[2]] + ggtitle("Uni. prior, big data - Daina"),
    Mikkel_uf[[2]] + ggtitle("Uni. prior, big data - Mikkel")
  )
uniform_teachers

# Uniform prior and regular amount of data
dens = 101 
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- rep(1 , dens)

Riccardo_uf_reg <- FindMyPosteriors(correct = 3,total = 6, prior = prior, p_grid = p_grid)
Riccardo_uf_reg[[1]]
Riccardo_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Riccardo") 

Tylén_uf_reg <- FindMyPosteriors(correct = 2, total = 2, prior = prior, p_grid = p_grid)
Tylén_uf_reg[[1]] 
Tylén_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Tylén")

Daina_uf_reg <- FindMyPosteriors(correct = 160, total = 198, prior = prior, p_grid = p_grid)
Daina_uf_reg[[1]] 
Daina_uf_reg[[2]]+ggtitle("Uniform prior,  reg. data - Daina") 

Mikkel_uf_reg <- FindMyPosteriors(correct = 66, total = 132, prior = prior, p_grid = p_grid)
Mikkel_uf_reg[[1]] 
Mikkel_uf_reg[[2]]+ggtitle("Uniform prior, reg. data - Mikkel")

uniform_reg_teachers <-
  cowplot::plot_grid(
    Riccardo_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Riccardo"),
    Tylén_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Tylén"),
    Daina_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Daina"),
    Mikkel_uf_reg[[2]] + ggtitle("Uni. prior, reg. data - Mikkel")
  )
uniform_reg_teachers

# Compare
cowplot::plot_grid(uniform_teachers,uniform_reg_teachers)
```
5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
*THOUGHTS:* + *i havent done any posterior predictive checks*
Setting a prior equal to 0? If our prior is that 0% knowledge is probable. We will give our "golem" the information state of 0 before the data. An initial confidence in each possible value of p at 0. 

Can be done like: 

OR maybe just not setting any prior

Setting a prior means that we give the "golem" an information state before the data. We will say that values of 0 are more plausible than 1's. We could do this by saying we want more values below 50 e.g.? Yes? 

We want a prior that says there is higher probability of teacher knowing 0% than 100%. Before, where we set a prior to numbers between 0-1 we said that the probability of the knowledge of CogSci was equal for all values of p. In other words, we did not know anything about their knowledge prior to the data collection. 

prior = dnorm(p_grid, 0.8,0.2) was used to set their knowledge at some point...

We could assign zero probability to all values over 0.5.
```{r}
# Uniform prior of 0 knowledge below 50%. 
dens = 202
p_grid <- seq(from = 0 , to = 1 , length.out = dens)
prior <- ifelse(p_grid > 0.5, 0, 1)

YouKnowNothingRiccardoSnow <- FindMyPosteriors(correct = 3,total = 6, prior = prior, p_grid = p_grid)
YouKnowNothingRiccardoSnow[[1]]
YouKnowNothingRiccardoSnow[[2]]+ggtitle("Uniform prior of 0 - Teacher background check") 

YouKnowNothingRiccardoSnow <- FindMyPosteriors(correct = 300,total = 600, prior = prior, p_grid = p_grid)
YouKnowNothingRiccardoSnow[[1]]
YouKnowNothingRiccardoSnow[[2]]+ggtitle("Uniform prior of 0 - Teacher background check - more data") 

# Trying something else 
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)? 
```{r}

```

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Daina: 148 correct answers out of 172 questions (again, Daina never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)



